\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-09-19 demystify.tex
various updates

\newenvironment{commentbox}[1][]{
\small
    \begin{cbox}
    \textbf{#1}: 
 }{
   \end{cbox}
}

@@

\section{Should You Apply?}
\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{The Simpsons}

@@

\begin{commentbox}[Vu]
One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam). In general the number of
PhD applications from Vietnam to US universities is low and more would be very welcomed. 
\end{commentbox}

@@

\epigraph{"How is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some of the old stuff out of my brain. Remember when I took that home wine making course, and I forgot how to drive?"}{The Simpsons}

@@

\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{The Simpsons}

@@

\epigraph{"Oh... and how is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some old stuff out of my brain. Remember when I took that home wine-making course and I forgot how to drive?"}{The Simpsons}

@@

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{The Simpson}

@@

\section{Miscs and FAQs}
\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{The Simpsons}

@@

\begin{commentbox}[Vu]
In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application. However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.

\tcblower
I would also advocate for that student as this shows they have the persistence and determination required for research.
\end{commentbox}

@@

\begin{commentbox}[Good example]

Dear Prof. Nguyen,

@@

...
\tcblower
This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G). 
\end{commentbox}

@@

\subsection{PhD in other related field such as CE, IST, Cybersecurity}

In many cases you do not need to do a PhD in CS to work in your area of interested. For example, in addition to having a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is very common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.

@@

If you don't know the professor (e.g., first email contact), then use \textbf{Prof. Lastname} or \textbf{Dr. Lastname}. I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.  Writing like that makes it like you copy and paste names, so no need to do so,  just Prof. or Dr. Lastname.
Moreover, if you don't know the prof., do not use Mr. or Mrs., or Firstname. To me it seems a bit disrespectful. As you know that prof. better and depends on their preference, you may call them by their Firstname.

Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr. Lastname.  So just observe and follow how it is done at your place.

\begin{commentbox}[Vu]
Dr. Vu is a bit amusing but I am totally fine with it.
\end{commentbox}

@@

\paragraph{Diversity} US universities prioritize diversity and inclusion. It's crucial for students to respect and appreciate varied opinions, backgrounds, and cultures. Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities). Racism or discrimination can have serious consequences, including academic and disciplinary actions.

\paragraph{Academic Integrity} Plagiarism and cheating, e.g., exams, coding, is a big no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is very different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will completely take over and investigate the case; and finally (iii) OAI make the final decision.  It is important to note that after receiving the report from your prof., OAI completely takes over (so begging your professor will not help because they simply are no longer involved).

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries. But do not install or use them on university computer (e.g., computers given by your department or advisor).  It is unlikely that the university will track you down, but it is the \textbf{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.   If you need software, ask your advisor or the department.

>>>>

\newenvironment{commentbox}[1][]{
\small
    \begin{cbox}
    \textbf{#1} 
 }{
   \end{cbox}
}

@@

\section{Should You Apply?}
\epigraph{\vspace{-0.2in} Don't make fun of graduate students. They just made a terrible life choice.}{\textsc{The Simpsons}}

@@

\begin{commentbox}[Vu]
One of the reasons I create this document is that my colleagues at GMU are interested in recruiting Vietnamese students and are surprised when seeing very few applications in Vietnam (e.g., each year our CS program receives more than 350 PhD applications, most of which are international but only 3--4 are from Vietnam). In general the number of CS PhD applications from Vietnam to US universities is low and more would be very welcomed. 
\end{commentbox}

@@

\epigraph{How is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some of the old stuff out of my brain. Remember when I took that home wine making course, and I forgot how to drive?}{\textsc{The Simpsons}}

@@

\epigraph{Son, if you really want something in this life, you have to work for it. Now quiet! They’re about to announce the lottery numbers.}{\textsc{The Simpsons}}

@@

\epigraph{"Oh... and how is education supposed to make me feel smarter? Besides, every time I learn something new, it pushes some old stuff out of my brain. Remember when I took that home wine-making course and I forgot how to drive?"}{\textsc{The Simpsons}}

@@

\epigraph{Bart, with \$10,000, we’d be millionaires! We could buy all kinds of useful things like…love!’}{\textsc{The Simpsons}}

@@

\section{Miscs and FAQs}
\epigraph{"I want to share something with you – the three little sentences that will get you through life; number 1: Cover for me, number 2: Oh, good idea, Boss, and number 3: It was like that when I got here."}{\textsc{The Simpsons}}

@@

In his \href{https://matt.might.net/articles/how-to-apply-and-get-in-to-graduate-school-in-science-mathematics-engineering-or-computer-science/}{post}, Matt Might was initially unsure about an application. However, upon learning that the applicant had led a 100km hike in the Himalayas, he decided to accept the applicant.  This is a good example of being "stand out", and I would also advocate for that student as this shows they have the persistence and determination required for research.

@@

\begin{commentbox}[Dear Prof. Nguyen,]

@@

...

This is a good example because it is clearly written just for me.  It shows that the student knows about my work on invariant generation and has related  background (paper C and project G). 
\end{commentbox}

@@

\subsection{PhD in other related fields: CE, IST, Cybersecurity}

In many cases you do not have to do a PhD in CS to work in your area of interested. For example, in addition to a traditional CS department, GMU has IST and Cybersecurity departments, which have faculty  with PhD in CS and work on CS topics (e.g., AI, Security, Robotics).  So it is totally possible that you still get to do CS research and publish in CS-focused venues even if you're not in a traditional CS program.  It is  common to see faculty with CS PhD in a non-CS department as well as faculty with non-CS PhD in CS department.

@@

If you don't know the professor (e.g., first email contact), then use \textbf{Prof. Lastname} or \textbf{Dr. Lastname}. I've seen many international students write \textbf{Prof.} or \textbf{Dr.} \textbf{Firstname Lastname}.  Writing like that makes it like you copy and paste names, so no need to do so,  just Prof. or Dr. Lastname.
Moreover, if you don't know the prof., do not use Mr. or Mrs., or Firstname. To me it seems a bit disrespectful. As you know that prof. better and depends on their preference, you may call them by their firstname. Many students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it (and Vu is way easier to pronounce than Nguyen).

Note that at some universities, especially those that do not offer PhD studies, will use the formal \textbf{Dr.} title, i.e., Dr. Lastname.  So just observe and follow how it is done at your place.  Also note that some faculty might not have a Ph.D., so they might not be Dr.  In that case, you can use Prof. Lastname.

@@

\paragraph{Diversity} US universities prioritize diversity and inclusion. Students need to respect and appreciate varied opinions, backgrounds, and cultures. Unlike some countries where certain voices are marginalized, in the US, all perspectives are valued equally (especially at universities). Racism or discrimination will have serious consequences, including academic and disciplinary actions.

\paragraph{Academic Integrity} Plagiarism and cheating (e.g., exams, coding) is a big no-no in the US.  If you're caught cheating, you will face very heavy consequence and likely be expelled from the university (e.g., after the second time at GMU).   This is very different from many international countries where cheating is common and often tolerated.  Faculty is extremely good at detecting cheating (we have been dealing with these situations so many times over so many years), and \emph{will} report cheating cases.  In short, whatever you do, don't cheat---not worth it.

Here is the typically steps: (i) a faculty suspecting a cheating case will report it to the Office of Academic Integrity (OAI) at the university;  (ii)  OAI will take over and investigate the case; and finally (iii) OAI will make the final decision.  It is important to note that after receiving the report from your prof., OAI \emph{completely} takes over (so begging your professor will not help because they simply are no longer involved and can't do much here).

\paragraph{Illegal Software} Using illegal/cracked software is very common in many countries. But do not install or use them on university computer (e.g., computers given by your department or advisor).  It is unlikely that the university will track you down, but it is the \emph{software company} that will.  They have very sophisticated tools to detect illegal software and will sue your university/department.  Imagine your department or advisor being sued for half a million dollars, and it is you who caused it.   If you need software, ask your advisor or the department.
